Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-10-04 writeup/intro.tex
More tweaks
more writing

At a high level, the virtues of HATs are guaranteed responses from any
replica, low latency, and a range of semantic guarantees including
several whose usefulness is widely acepted such as Read
Committed. However, highly available systems are fundamentally unable
to prevent concurrent updates to shared data items and cannot provide
recency guarantees for reads. To understand when these virtues and
limitations are relevant in practice, we survey both practitioner
accounts and academic literature, perform experimental analysis on
modern cloud infrastructure, and analyze representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints for multi-item updates and, in some cases, provide limited
uniqueness guarantees. However, particularly for programs with
non-monotonic logic or that are otherwise concurrency-sensitive, HATs
can fall short, requiring non-HA coordination.

We recognize that the large variety of ACID isolation levels and
distributed consistency models (and therefore in our taxonomization)
can be confusing and that the subtle distinctions between them can
often appear to be primarily of academic concern. Accordingly, we
offer the following pragmatic takeaways:
\begin{introenumerate}
\item The default (and sometimes strongest) configurations of most
  widely-deployed database systems expose a range of anomalies that
  can compromise application-level consistency.

\item Most of these ``weak isolation'' models are achievable without
  sacrificing high availability if implemented correctly. However,
  none of these models prevents concurrent modifications.

>>>>

At a high level, the virtues of HATs are guaranteed responses from any
replica, low latency, and a range of semantic guarantees including
several whose usefulness is widely acepted such as Read
Committed. However, highly available systems are fundamentally unable
to prevent concurrent updates to shared data items and cannot provide
recency guarantees for reads. To understand when these virtues and
limitations are relevant in practice, we survey both practitioner
accounts and academic literature, perform experimental analysis on
modern cloud infrastructure, and analyze representative applications
for their semantic requirements. Our experiences with a HAT prototype
running across multiple geo-replicated datacenters indicate that HATs
offer a one to three order of magnitude latency decrease compared to
traditional distributed serializability protocols, and they can
provide acceptable semantics for a wide range of programs, especially
those with monotonic logic and commutative updates~\cite{calm,
  crdt}. HAT systems can also enforce arbitrary foreign key
constraints for multi-item updates and, in some cases, provide limited
uniqueness guarantees. However, HATs can fall short for applications
with concurrency-sensitive operations, requiring non-HA coordination.

Finally, we recognize that the large variety of ACID isolation levels
and distributed consistency models (and therefore those in our
taxonomization) can be confusing; the subtle distinctions between
models can appear to be of primarily academic concern. Accordingly, we
offer the following pragmatic takeaways:
\begin{introenumerate}
\item The default (and sometimes strongest) configurations of most
  widely-deployed database systems expose a range of anomalies that
  can compromise application-level consistency.

\item Many of these ``weak isolation'' models are achievable without
  sacrificing high availability if implemented correctly. However,
  none of these models prevents concurrent modifications.

### 2013-10-04 writeup/motivation.tex
more writing

These results---which do not consider server failures---indicate
that network partitions \textit{do} occur within and across modern
datacenters. We observe that, as a general trend, partitions exist and
must accordingly be met with either unavailability at some servers or,
as we will discuss, relaxed semantic guarantees.

@@

\caption{Mean RTT times on EC2 (min and max highlighted)}\vspace{-.5em}
\label{table:rtt}
\end{table}

>>>>

These studies indicate that network partitions \textit{do} occur
within and across modern datacenters. We observe that these partitions
must be met with either unavailability at some servers or, as we will
discuss, relaxed semantic guarantees.

@@

\caption{Mean RTT times on EC2 (min and max highlighted)}\vspace{-1em}
\label{table:rtt}
\end{table}
